Think deep for reef resilience

### 2025-07-11 paper/paper_draft.qmd
Comments for @eco-ben

Statistically downscaled 1km resolution Sea Surface Temperature (SST) data for the Great
Barrier Reef were obtained from the HighResCoralStress database [@dixon2022; @dixon].
Maximum monthly mean DHW levels were calculated from the SST data following NOAA methodology
[@liu2017; @skirving2020]. The SSTs are represented by a selection of five Global Climate
Models (GCMs) from the Sixth Phase of the Coupled Model Intercomparison Project (CMIP6;
**citation**). These five GCMs were EC-Earth3-Veg, ACCESS-CM2, ACCESS-ESM1-5, NorESM2-MM and
GDFL-CM4, and their inclusion were based on evaluation by Grose et al. [-@grose2023], which
found high correlation between observed temperature, rainfall, El Niño Southern Oscillation
and Indian Ocean Dipole patterns in the Australian terrestrial region, and the corresponding
mechanisms in these GCMs.

Equilibrium climate sensitivity (ECS) is defined, by the IPCC, as the long term warming
caused by a doubling of carbon dioxide above its pre-industrial concentration
[@forster2021]. The four GCMs span a range of ECS values within the IPCC assessed likely
range, and GCM development families, allowing for a range of possible climate futures to be
explored, capturing the uncertainty in future responses to CO2 increases (@fig-ecs;
[@grose2023; @shepherd2018]). The following analyses are performed using each GCM
separately. SSP 2-4.5 is used as a plausible emission pathway under current emission
trajectories, with similar atmospheric CO2 levels remaining until 2050 [@hausfather2020;
IPCC -@climate2023].

>>>>

Statistically downscaled 1km resolution Sea Surface Temperature (SST) data for the Great
Barrier Reef were obtained from the HighResCoralStress database [@dixon2022; @dixon].
Maximum monthly mean DHW levels were calculated from the SST data following NOAA methodology
[@liu2017; @skirving2020]. The SSTs are represented by a selection of five Global Climate
Models (GCMs) from the Sixth Phase of the Coupled Model Intercomparison Project (CMIP6;
**citation**). These five GCMs were EC-Earth3-Veg, ACCESS-CM2, ACCESS-ESM1-5, NorESM2-MM and
GDFL-CM4, and their inclusion were based on evaluation by Grose et al. [-@grose2023], which
found high correlation between observed temperature, rainfall, El Niño Southern Oscillation
and Indian Ocean Dipole patterns in the Australian terrestrial region, and the corresponding
mechanisms in these GCMs.
<!-- @Ben, the last sentence above is quite long and doesn't explain why the five GCMs
were selected. -->

Equilibrium climate sensitivity (ECS) is defined, by the IPCC, as the long term warming
caused by a doubling of carbon dioxide above its pre-industrial concentration
[@forster2021]. The four GCMs
<!-- @Ben, says "four GCMs" here, but we used five? -->
span a range of ECS values within the IPCC assessed likely
range, and GCM development families, allowing for a range of possible climate futures to be
explored, capturing the uncertainty in future responses to CO2 increases (@fig-ecs;
[@grose2023; @shepherd2018]). The following analyses are performed using each GCM
separately. SSP 2-4.5 was
<!-- I changed "is" to "was" - anything to do with the analysis making the decision to
standardise to past tense as currently there is a mix. -->
used as a plausible emission pathway under current emission
trajectories, with similar atmospheric CO2 levels remaining until 2050 [@hausfather2020;
IPCC -@climate2023].
